The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2018-08-08 script/en/chapters/cryptomethods.tex
CT book (de+en): Corrected all broken links and fixed some few typos.

\vskip +3pt
With the use of the internet and wireless communication, encryption technologies
are used (mostly transparently) by everyone. However, they have been in use since
centuries by governments, military, and diplomats. The side which had a better command
of these technologies could exert big influence on {\bf politics} and war with the help
of secret services.
This book only touches history when introducing the earlier cipher methods in
chapter \ref{Chapter_PaperandPencil}. You can gain an impression, how important
cryptology was and still is, by considering the following two examples: the
educational film ``War of the letters'' (German: Krieg der Buchstaben)\footnote{%
  See \url{http://bscw.schule.de/pub/bscw.cgi/d1269787/Krieg_der_Buchstaben.pdf}.\\
  A supporting quote from Denis Smyth, professor in the Department of History
  and the International Relations Programme at the University of Toronto:
  ``Secret intelligence has long been regarded as the ``missing dimension'' of
  international relations.'' (from \url{http://www.secretintelligencefiles.com})
}
and the debates around the so called crypto wars\footnote{%
  See \url{https://en.wikipedia.org/wiki/Crypto_Wars}.\index{crypto wars}
}.

@@

The Reference for SageMath v7.2 lists e.g. the following further cryptographic functions:%
\footnote{
  See
  \url{http://doc.sagemath.org/html/en/reference/sage/crypto/index.html},\\
  \url{http://doc.sagemath.org/html/en/reference/cryptography/index.html}, and\\
  \url{http://combinat.sagemath.org/doc/reference/cryptography/sage/crypto/stream.html}
}
\begin{itemize}
   \item Linear feedback shift register (LFSR),
   \item Blum-Blum-Shub (BBS): pseudo-random generator (to be found with streams),
   \item Lattice-based functions.
\end{itemize}

@@

  \hypertarget{CM_HT_Weblink_Rijndael-Cryptosystem}{}
  \item AES discussion groups at NIST (archive page provided for historical purposes, last update on Feb 28th, 2001)\\
	\url{http://csrc.nist.gov/archive/aes/}

@@

All links have been confirmed at July 10, 2016.

>>>>

\vskip +3pt
With the use of the internet and wireless communication, encryption technologies
are used (mostly transparently) by everyone. However, they have been in use since
centuries by governments, military, and diplomats. The side which had a better command
of these technologies could exert big influence on {\bf politics} and war with the help
of secret services.
This book only touches history when introducing the earlier cipher methods in
chapter \ref{Chapter_PaperandPencil}. You can gain an impression, how important
cryptology was and still is, by considering the following two examples: the
educational film ``War of the letters'' (German: Krieg der Buchstaben)\footnote{%
  See \url{https://bscw.schule.de/pub/bscw.cgi/d1269787/Krieg_der_Buchstaben.pdf}.\\
  A supporting quote from Denis Smyth, professor in the Department of History
  and the International Relations Programme at the University of Toronto:
  ``Secret intelligence has long been regarded as the ``missing dimension'' of
  international relations.'' (from \url{http://www.secretintelligencefiles.com})
}
and the debates around the so called crypto wars\footnote{%
  See \url{https://en.wikipedia.org/wiki/Crypto_Wars}.\index{crypto wars}
}.

@@

The Reference for SageMath v8.2 lists e.g. the following further cryptographic functions:%
\footnote{
  See
  \url{http://doc.sagemath.org/html/en/reference/sage/crypto/index.html},\\
  \url{http://doc.sagemath.org/html/en/reference/cryptography/index.html}, and\\
  \url{http://doc.sagemath.org/html/en/reference/cryptography/sage/crypto/stream.html}
}
\begin{itemize}
   \item Linear feedback shift register (LFSR),
   \item Blum-Blum-Shub (BBS): pseudo-random generator (to be found with streams),
   \item Lattice-based functions.
\end{itemize}

@@

  \hypertarget{CM_HT_Weblink_Rijndael-Cryptosystem}{}
  \item AES discussion groups at NIST (archive page provided for historical purposes, last update on Feb 28th, 2001)\\
	\url{https://csrc.nist.gov/archive/aes/}

@@

All links have been confirmed at August 6, 2018.

### 2018-08-08 script/en/chapters/digitalsignatures.tex
CT book (de+en): Corrected all broken links and fixed some few typos.

To address new findings in cryptanalysis, in 2008 NIST opened a competition to
develop a new cryptographic hash algorithm beyound the SHA-2 family: In October
2012 Keccak\index{Keccak}\index{SHA-3} was announced as ``SHA-3''.\footnote{%
\url{http://csrc.nist.gov/groups/ST/hash/sha-3/}\\
    With {\bf CT2}\index{CT2} you can execute and visualize
    the Keccak hash function -- using in the Startcenter
    {\bf Templates \textbackslash{} Hash Functions \textbackslash{}
    Keccak Hash (SHA-3)}.\\
    Keccak also can be used as pseudo random number generator and as stream
    cipher: This can be found within the Startcenter templates via
    {\bf Tools \textbackslash{} Keccak PRNG}, and 
    {\bf Cryptography \textbackslash{} Modern Ciphers \textbackslash{}
         Symmetric \textbackslash{} Keccak Streamcipher}.
}

Further information about this topic can be found in the article
``Hash cracked -- The consequences of the successful attacks on SHA-1''
by Reinhard Wobst and J\"urgen Schmidt\footnote{% 
      \url{http://www.heise.de/security/artikel/56634}.\\
  Further references are e.g.:\\
      \url{http://csrc.nist.gov/groups/ST/toolkit/index.htmll}.
} 
  by Heise Security.

@@

\noindent All links have been confirmed at July 14, 2016.

>>>>

To address new findings in cryptanalysis, in 2008 NIST opened a competition to
develop a new cryptographic hash algorithm beyound the SHA-2 family: In October
2012 Keccak\index{Keccak}\index{SHA-3} was announced as ``SHA-3''.\footnote{%
\url{https://csrc.nist.gov/groups/ST/hash/sha-3/}\\
    With {\bf CT2}\index{CT2} you can execute and visualize
    the Keccak hash function -- using in the Startcenter
    {\bf Templates \textbackslash{} Hash Functions \textbackslash{}
    Keccak Hash (SHA-3)}.\\
    Keccak also can be used as pseudo random number generator and as stream
    cipher: This can be found within the Startcenter templates via
    {\bf Tools \textbackslash{} Keccak PRNG}, and 
    {\bf Cryptography \textbackslash{} Modern Ciphers \textbackslash{}
         Symmetric \textbackslash{} Keccak Streamcipher}.
}

%% BE_8.8.18: Englische Fassung von heise.de nicht mehr erreichbar.
%% Zumindest nicht die bisher angegebene Nummer.
%% ``Hash cracked -- The consequences of the successful attacks on SHA-1''
%%      \url{http://www.heise.de/security/artikel/56634}.\\
Further information about this topic can be found in the article
``Consequences of the successful attacks on SHA-1''
by Reinhard Wobst and J\"urgen Schmidt\footnote{% 
      \url{https://www.heise.de/security/artikel/Consequences-of-the-successful-attacks-on-SHA-1-270650.html}.\\
  Further references are e.g.:\\
      \url{https://csrc.nist.gov/groups/ST/toolkit/index.htmll}.
} 
  by Heise Security.

@@

\noindent All links have been confirmed at August 6, 2018.

### 2018-08-08 script/en/chapters/ellipticcurves.tex
CT book (de+en): Corrected all broken links and fixed some few typos.

If we assume that the computing power increases by Moore's law (i.~e.\ it
doubles every 18 months)\footnote{%
  Moore's law formulates the empirical observation and the according
  forecast that the number of components or transistors on an integrated circuit
  doubles every two years. It referred originally only to the transistors density
  in an integrated circuit, however not, for example, to the increase in the storage
  density.
  This realization for transistors density in 1965 (with a correction 1975) was
  expressed by Gordon Moore, a co-founder of Intel.
  In recent years, the growth of computing power was even higher than
  the forecasted doubling every 2 years. Future limits are set by the
  transistors size of a few atoms, which could be achieved by 2020.
}% https://en.wikipedia.org/wiki/Moore%27s_law
, then the evolution of the key lengths for secure communication will be as
in figure~\ref{RSAKeylength}, which was generated from table 1 (on page 32 in
\cite{LeVe2001}).\footnote{%
Further information about key length comparison by Arjen Lenstra und Eric
Verheul, plus more modern evaluations till 2015 can be found in the
interactive website \url{http://www.keylength.com}.
}

@@

\begin{sloppypar}
\begin{itemize}
    \item \url{http://doc.sagemath.org/html/en/constructions/elliptic_curves.html}
    \item \url{http://doc.sagemath.org/html/en/reference/plane_curves/index.html#elliptic-curves}
\end{itemize}
\end{sloppypar}

@@

   \item Working group IEEE P1363,\\
        \url{http://grouper.ieee.org/groups/1363}

@@

>>>>

If we assume that the computing power increases by Moore's law (i.~e.\ it
doubles every 18 months)\footnote{%
  Moore's law formulates the empirical observation and the according
  forecast that the number of components or transistors on an integrated circuit
  doubles every two years. It referred originally only to the transistors density
  in an integrated circuit, however not, for example, to the increase in the storage
  density.
  This realization for transistors density in 1965 (with a correction 1975) was
  expressed by Gordon Moore, a co-founder of Intel.
  In recent years, the growth of computing power was even higher than
  the forecasted doubling every 2 years. Future limits are set by the
  transistors size of a few atoms, which could be achieved by 2020.
}% https://en.wikipedia.org/wiki/Moore%27s_law
, then the evolution of the key lengths for secure communication will be as
in figure~\ref{RSAKeylength}, which was generated from table 1 (on page 32 in
\cite{LeVe2001}).\footnote{%
Further information about key length comparison by Arjen Lenstra und Eric
Verheul, plus more modern evaluations till 2015 can be found in the
interactive website \url{https://www.keylength.com}.
}

@@

\begin{sloppypar}
\begin{itemize}
    \item \url{http://doc.sagemath.org/html/en/constructions/elliptic_curves.html}
    \item \url{http://doc.sagemath.org/html/en/reference/curves/index.html#elliptic-curves}
\end{itemize}
\end{sloppypar}

@@

   \item Working group IEEE P1363,\\
		IEEE P1363 is an IEEE standardization project for public-key cryptography.\\
		Elliptic curves were part of so called established or traditional public-key documents. The project began in 1993. The group's original website \textit{http://grouper.ieee.org/groups/1363} is no more accessible (status of 2018-08-08).

@@

\noindent All links have been confirmed at August 6, 2018.

### 2018-08-08 script/en/chapters/moderncryptography.tex
CT book (de+en): Corrected all broken links and fixed some few typos.

The current record\index{logarithm problem!record} (as of April 2007) for calculating discrete logarithms was
established in February 2007 by the group Kleinjung, Franke, and Bahr at
University of Bonn.\footnote{%
   \url{http://www.nabble.com/Discrete-logarithms-in-GF(p)-----160-digits-t3175622.html}}
Kleinjung calculated the discrete logarithm modulo a 160 digit prime number $p$
and generator $g$:
$$
\begin{array}{r@{\:}c@{\:}l}
p & = & \lfloor 10^{159}\pi\rfloor + 119849\\
  & = & 314159265358979323846264338327950288419716939937510582097494\\
  &   & 459230781640628620899862803482534211706798214808651328230664\\
  &   & 7093844609550582231725359408128481237299\\
g & = & 2
\end{array}
$$
The discrete logarithms $k$ of the following integer $y$ was
determined:\footnote{The integer $y$ was chosen as the first 159 digits of the
Euler number $e$.}
$$
\begin{array}{r@{\:}c@{\:}l}
y & = & \lfloor 10^{159}e\rfloor\\
  & = & 271828182845904523536028747135266249775724709369995957496696\\
  &   & 762772407663035354759457138217852516642742746639193200305992\\
  &   & 1817413596629043572900334295260595630738\\
k & = & \log_g(y)\mod p\\
  & = & 829897164650348970518646802640757844024961469323126472198531\\
&   & 845186895984026448342666252850466126881437617381653942624307\\
&   & 537679319636711561053526082423513665596
\end{array}
$$
The search was performed with GNFS method (general number field sieve,
Index-Calculus) \index{general number field sieve (GNFS)} and took about
17 CPU years on 3.2 GHz Xeon machines.

>>>>

% BE_2018: TODOTODO: Aktualisieren !
% The current record\index{logarithm problem!record} (as of April 2007) for calculating discrete logarithms was
% established in February 2007 by the group Kleinjung, Franke, and Bahr at
% University of Bonn.
In February 2007, the group Kleinjung, Franke, and Bahr at University of Bonn
did the then record\index{logarithm problem!record} for calculating discrete logarithms.
They calculated the discrete logarithm modulo a 160 digit prime number $p$
and a generator $g$:
$$
\begin{array}{r@{\:}c@{\:}l}
p & = & \lfloor 10^{159}\pi\rfloor + 119849\\
  & = & 314159265358979323846264338327950288419716939937510582097494\\
  &   & 459230781640628620899862803482534211706798214808651328230664\\
  &   & 7093844609550582231725359408128481237299\\
g & = & 2
\end{array}
$$
The discrete logarithms $k$ of the following integer $y$ was
determined:\footnote{The integer $y$ was chosen as the first 159 digits of the
Euler number $e$.}
$$
\begin{array}{r@{\:}c@{\:}l}
y & = & \lfloor 10^{159}e\rfloor\\
  & = & 271828182845904523536028747135266249775724709369995957496696\\
  &   & 762772407663035354759457138217852516642742746639193200305992\\
  &   & 1817413596629043572900334295260595630738\\
k & = & \log_g(y)\mod p\\
  & = & 829897164650348970518646802640757844024961469323126472198531\\
&   & 845186895984026448342666252850466126881437617381653942624307\\
&   & 537679319636711561053526082423513665596
\end{array}
$$
The search was performed with GNFS method (general number field sieve,
Index-Calculus) \index{general number field sieve (GNFS)} and took about
17 CPU years on 3.2 GHz Xeon machines.

### 2018-08-08 script/en/chapters/movies.tex
CT1: New menutree pdfs: Updated CTB and copied CTB pdfs in setup\template-xx; corrected according online help files hlp_xx/menu_structure.html
CT book (de+en): Corrected all broken links and fixed some few typos.

\item[\textrm{[Kipling1901]}] \index{Kipling 1901}
    Rudyard Kipling\index{Kipling, Rudyard}, \\
    {\em Kim}, 1901. \\
    Rob Slade's review%
    \footnote{See
        \url{http://catless.ncl.ac.uk/Risks/24.49.html#subj12}.
    }
    of this novel says:
    ``Kipling packed a great deal of information and concept into his stories,
    and in ``Kim'' we find The Great Game: espionage and spying.  Within the
    first twenty pages we have authentication by something you have, denial
    of service, impersonation, stealth, masquerade, role-based
    authorization (with ad hoc authentication by something you know),
    eavesdropping, and trust based on data integrity.  Later on we get
    contingency planning against theft and cryptography with key changes.''\\
    The book is out of copyright%
    \footnote{You can read it at:\\
          \url{http://whitewolf.newcastle.edu.au/words/authors/K/KiplingRudyard/prose/Kim/index.html},\\
          \url{http://kipling.thefreelibrary.com/Kim} or\\
          \url{http://www.readprint.com/work-935/Rudyard-Kipling}.
    }.\\

@@

\item[\textrm{[Robinson1992]}] \index{Robinson 1992}
    Directed by Phil Alden Robinson, \\
    {\em Sneakers}, Universal Pictures Film, 1992. \\
    In this movie the ``sneakers'', computer experts under their boss Martin
    Bishop, try to get back the deciphering box SETEC from the ``bad guys''.
    SETEC, invented by a genius mathematician before he was killed, allows to
    decrypt all codes from any nation. \\
    In the movie the code is not described in any way%
    \footnote{
       Leonard Adleman (the ``A'' within RSA) worked as mathematical consultant
       for ``Sneakers''. He describes the funny story about his contribution
       at his homepage 
       \url{http://www.usc.edu/dept/molecular-science/fm-sneakers.htm}.
       It is assumed that the cipher used everywhere is RSA.      
       According to that within the chip a fast, unknown factorization
       method\index{factorization} is implemented.
    }.\\

>>>>

\item[\textrm{[Kipling1901]}] \index{Kipling 1901}
    Rudyard Kipling\index{Kipling, Rudyard}, \\
    {\em Kim}, 1901. \\
    Rob Slade's review%
    \footnote{See
        \url{http://catless.ncl.ac.uk/Risks/24.49.html#subj12}.
    }
    of this novel says:
    ``Kipling packed a great deal of information and concept into his stories,
    and in ``Kim'' we find The Great Game: espionage and spying.  Within the
    first twenty pages we have authentication by something you have, denial
    of service, impersonation, stealth, masquerade, role-based
    authorization (with ad hoc authentication by something you know),
    eavesdropping, and trust based on data integrity.  Later on we get
    contingency planning against theft and cryptography with key changes.''\\
    The book is out of copyright.%
    \footnote{You can read it at:\\
          \url{http://whitewolf.newcastle.edu.au/words/authors/K/KiplingRudyard/prose/Kim/index.html},\\
          \url{http://kipling.thefreelibrary.com/Kim} or\\
          \url{http://www.readprint.com/work-935/Rudyard-Kipling}.
    }\\

@@

\item[\textrm{[Robinson1992]}] \index{Robinson 1992}
    Directed by Phil Alden Robinson, \\
    {\em Sneakers}, Universal Pictures Film, 1992. \\
    In this movie the ``sneakers'', computer experts under their boss Martin
    Bishop, try to get back the deciphering box SETEC from the ``bad guys''.
    SETEC, invented by a genius mathematician before he was killed, allows to
    decrypt all codes from any nation. \\
    In the movie the code is not described in any way%
    \footnote{
       Leonard Adleman (the ``A'' within RSA) worked as mathematical consultant
       for ``Sneakers''. He describes the funny story about his contribution
       at his homepage 
       \url{http://world.std.com/~reinhold/math/sneakers.adleman.html}.
       It is assumed that the cipher used everywhere is RSA.      
       According to that within the chip a fast, unknown factorization
       method\index{factorization} is implemented.
    }.\\

### 2018-08-08 script/en/chapters/numbertheory.tex
CT book (de+en): Corrected all broken links and fixed some few typos.

The 6 smaller prime factors, already found before have been:
$$
\begin{array}{c}
3, 1907, 425796183929,\\
1624700279478894385598779655842584377,\\
3802306738549441324432139091271828121 \quad {\rm and}\\
128064886830166671444802576129115872060027.
\end{array}
$$
The first 3 factors can be easily computed\footnote{%
E.g.\ using CT1\index{CT1} via menu
{\bf Indiv. Procedures \textbackslash{} RSA Cryptosystem \textbackslash{}
Factorization of a Number}.\\
CT1 can factorize in a reasonable time numbers no longer than 250 bit (Numbers
bigger than 1024 bits are not accepted by CT1). CT2\index{CT2} is able
to factorize numbers bigger than 250 bit length.
}.
The next three prime factors were found by P.~Zimmerman\footnote{%
\url{http://homepages.loria.fr/PZimmermann/ecmnet/}},
T.~Grandlund and R. Harley during the years 1999 and 2000 using the elliptic curve
factorization method.

@@

The task of the researchers was not initiated by a challenge, but
they wanted to find the last prime factors of the Mersenne number $2^{1039}+1$
(see ``Wanted List'' of the Cunningham Project\index{Cunningham project}\footnote{%
Cunningham project: \url{http://homes.cerias.purdue.edu/~ssw/cun/}\\
Cunningham table: \url{http://homes.cerias.purdue.edu/~ssw/cun/pmain1215}\\
The numbers in the Cunningham table have the following syntax:\\
``(2,n)-'' means $2^{n}-1$;~~~
``(2,n)+'' means $2^{n}+1$.\\
To describe the magnitude one writes $p<n>$ or $c<n>$:
``n'' is the number of decimal digits and ``p'' and ``c'' tell,
whether the number is prime or composite.\\
$2^{1039}-1 = p7 * c307 = p7 * p80 * p227$\\
It is explained more precisely at the page of the Cunningham project:\\
``2651+ means $2^{651} + 1$ and the size (c209 means 209 decimal digits)
of the
number which was factored.  Then come the new factor(s), the discoverer and
the method used.  Recently, only the multiple polynomial quadratic sieve
(ppmpqs), the elliptic curve method (ecm) and the number field sieve (nfs)
have been used.  `hmpqs' stands for hypercube multiple polynomial quadratic
sieve.  Under `new factors', `p90' means a 90-digit prime and `c201' is a
201-digit composite number.''.
}).\\

>>>>

The 6 smaller prime factors, already found before have been:
$$
\begin{array}{c}
3, 1907, 425796183929,\\
1624700279478894385598779655842584377,\\
3802306738549441324432139091271828121 \quad {\rm and}\\
128064886830166671444802576129115872060027.
\end{array}
$$
The first 3 factors can be easily computed\footnote{%
E.g.\ using CT1\index{CT1} via menu
{\bf Indiv. Procedures \textbackslash{} RSA Cryptosystem \textbackslash{}
Factorization of a Number}.\\
CT1 can factorize in a reasonable time numbers no longer than 250 bit (Numbers
bigger than 1024 bits are not accepted by CT1). CT2\index{CT2} is able
to factorize numbers bigger than 250 bit length.
}.
The next three prime factors were found by P.~Zimmermann\footnote{%
\url{http://homepages.loria.fr/PZimmermann/ecmnet/}},
T.~Grandlund and R. Harley during the years 1999 and 2000 using the elliptic curve
factorization method.

@@

The task of the researchers was not initiated by a challenge, but
they wanted to find the last prime factors of the Mersenne number $2^{1039}+1$
(see ``Wanted List'' of the Cunningham Project\index{Cunningham project}\footnote{%
Cunningham project: \url{http://homes.cerias.purdue.edu/~ssw/cun/}\\
% BE_18-8-8: Link tot: Cunningham table: \url{http://homes.cerias.purdue.edu/~ssw/cun/pmain1215}\\
The numbers in the Cunningham table have the following syntax:\\
``(2,n)-'' means $2^{n}-1$;~~~
``(2,n)+'' means $2^{n}+1$.\\
To describe the magnitude one writes $p<n>$ or $c<n>$:
``n'' is the number of decimal digits and ``p'' and ``c'' tell,
whether the number is prime or composite.\\
$2^{1039}-1 = p7 * c307 = p7 * p80 * p227$\\
It is explained more precisely at the page of the Cunningham project:\\
``2651+ means $2^{651} + 1$ and the size (c209 means 209 decimal digits)
of the
number which was factored.  Then come the new factor(s), the discoverer and
the method used.  Recently, only the multiple polynomial quadratic sieve
(ppmpqs), the elliptic curve method (ecm) and the number field sieve (nfs)
have been used.  `hmpqs' stands for hypercube multiple polynomial quadratic
sieve.  Under `new factors', `p90' means a 90-digit prime and `c201' is a
201-digit composite number.''.
}).\\

### 2018-08-08 script/en/chapters/primes.tex
CT1: CT book (d+e): checked urls used within \href command and some cosmetics
CT book (de+en): Corrected all broken links and fixed some few typos.

\begin{figure}[!hb]
\begin{center}
\includegraphics[scale=0.3]{figures/Numberrectangle-with-colored-primes_Irisprime.jpg}
\caption[Primes within the first 390 integers -- marked with color]
        {Primes within the first 390 integers -- marked with color\protect\footnotemark}
\label{Primes-in-a-390-integer-rechtangle-figure}
\end{center}
\end{figure}
\footnotetext{%
   Graphics from 
   \url{http://mathforum.org/mathimages/index.php/Image:Irisprime.jpg}, 30*13 rectangle
   }

\begin{figure}[!hb]
\begin{center}
\includegraphics[scale=0.41]{figures/Ulam_spiral_via-CT2-within-1-999.png}
\caption[Primes within the first 999 integers -- as Ulam spiral]
        {Primes within the first 999 integers -- as Ulam spiral\protect\footnotemark}
\label{Primes-in-a-999-integer-ulam-spiral-figure}
\end{center}
\end{figure}
\footnotetext{%
   Graphics from CT2, menu Crypto Tutorials, World of Primes,
   Distribution of primes, Ulam's spiral; 32*32 points.\index{CT2}
}
%\newpage
\begin{figure}[!hb]
\begin{center}
\includegraphics{figures/Ulam_spiral_Wikipedia.png}
\caption[Primes within the first 4000 integers -- as Ulam spiral]
        {Primes within the first 4000 integers -- as Ulam spiral\protect\footnotemark}
\label{Primes-in-a-4000-integer-ulam-spiral-figure}
\end{center}
\end{figure}
\footnotetext{%
   Graphics from 
   \url{http://mathforum.org/mathimages/index.php/Image:Ulam_spiral.png}, 200*200 Ulam spiral
   }

@@

\hline
\end{tabular}
\caption{The 30+ largest known primes and its particular number types
         (as of Jan 2018)}    % Eyecatcher_neue-Mersenne
\label{L_n_Largest_Known-Primes}
\end{center}
\end{table} 
\addtocounter{footnote}{-1}
\footnotetext{\index{number!Sierpinski}\index{Seventeen or Bust SoB}%
This number was found within the distributed computing project
``Seventeen or Bust'' (SoB)  (\url{https://en.wikipedia.org/wiki/Seventeen_or_Bust})
at March 26, 2007. While the well known \hyperlink{GIMPS-project}{GIMPS project} (chapter~\ref{zahlentyp_mersenne}) searches for bigger and bigger of the infinitely many primes, there is a chance, that the SoB project could have been completed its task sometime.

@@

% --------------------------------------------------------------------------
\vskip +25 pt
\subsection{Challenge of the Electronic Frontier Foundation (EFF)}\index{EFF}
This search is also spurred on by a competition started by the non-profit
organization EFF (Electronic Frontier Foundation) using the means of an 
unknown donor. The participants are rewarded with a total of 500,000 USD if
they find the longest prime number. In promoting this project, the unknown
donor is not looking for the quickest computer, but rather wants to draw 
people's attention to the opportunities offered by {\em cooperative networking} \\
{\url{http://www.eff.org/awards/coop}}

@@

Considerable work has been done on Goldbach's weak conjecture, culminating in a 2013 claim by Harald Helfgott to fully prove the conjecture for all odd integers greater than 7.\footnote{%
Helfgott, H.A. (2013): ``Major arcs for Goldbach's theorem''.
\url{http://arxiv.org/abs/1305.2897}.\\
v4 of his 79 pages paper (14 Apr 2014) states, that the proof works starting at $10^{29}$ (instead of $10^{30})$. And for all numbers up to $10^{30}$ he claims, that he and David Platt proved at the computer that the weak Goldbach is valid.
\\
\url{http://truthiscool.com/prime-numbers-the-271-year-old-puzzle-resolved}.\\
\url{http://www.newscientist.com/article/dn23535-proof-that-an-infinite-number-of-primes-are-paired.html#.UgwhOpL0Ek0}.\\
\url{http://www.spiegel.de/wissenschaft/mensch/beweis-fuer-schwache-goldbachsche-vermutung-a-901111.html}.
}

@@

In May 2013 the results of Yitang Zhang\index{Zhang, Yitang} became
known.\footnote{%
    Erica Klarreich (May 19, 2013): ``Unheralded Mathematician Bridges the
    Prime Gap''\\
    \url{https://www.simonsfoundation.org/quanta/20130519-unheralded-mathematician-bridges-the-prime-gap/}}
Zhang proved, that there are infinitely many ``cousin primes'', or more
concretely that there is some number $H$ smaller than 70 million such that there
are inifinitely many pairs of primes that differ by $H$.\footnote{%
Whereas the gap between the primes of a twin prime is exactly $2$, cousin
primes\index{cousin prime} do denote two primes, which have a gap between them,
which has a value of a bigger, even, but finite number $H$.
      }${}^,$\footnote{%
This is close to the conjecture stated in 1849 by the French mathematician
Alphonse de Polignac that there are infinitely many prime pairs for any
possible even finite gap, not just $2$.
       }${}^,$\footnote{%
In the meantime this minimal gap $H$ of 70 millions was improved in further
work. The according progress is documented in the Polymath8 project (massively
collaborative online mathematical projects): ``Bounded gaps between primes''.
The best known value of $H$ was $4680$ (as of August 2013) and is till now
(as of April 2014) $2460$ -- this is a good progress compared to 70 million,
but far away from $2$.\\
See \url{http://michaelnielsen.org/polymath1/index.php?title=Bounded_gaps_between_primes}
   }

@@

\vskip +25 pt
\hypertarget{HT-GoogleRecruitment2004}{}
\subsubsection{Recruitment at Google in 2004}%
% \paragraph*{Recruitment at Google in 2004}%
\index{Google!recruitment}%
\label{HT-GoogleRecruitment2004}%
\mbox{}
In summer 2004 the company Google used the number $e$ to attract potential
employees.\footnote{The base of the natural logarithm\index{logarithm!natural}
$e$ is approximately 2.718 281 828 459.
This is one of the most important numbers in all of mathematics like complex
analysis, finance, physics and geometry. Now it was used the first time -- as
far as I know -- for marketing or recruitment.}${}^,$\footnote{Most of this
information is taken from the article ``e-number crunching'' by John Allen
Paulos in TheGuardian, Sept. 30, 2004, and from the web:\\
- \url{https://mkaz.tech/google-billboard-problems.html}\\
- \url{http://epramono.blogspot.com/2004/10/7427466391.html}\\
- \url{http://mathworld.wolfram.com/news/2004-10-13/google/}.
}

>>>>

\begin{figure}[!hb]
\begin{center}
\includegraphics[scale=0.3]{figures/Numberrectangle-with-colored-primes_Irisprime.jpg}
\caption[Primes within the first 390 integers -- marked with color]
        {Primes within the first 390 integers -- marked with color\protect\footnotemark}
\label{Primes-in-a-390-integer-rechtangle-figure}
\end{center}
\end{figure}
\footnotetext{%
   Graphics from 
   \url{http://mathforum.org/mathimages/index.php/Image:Irisprime.jpg}, 30*13 rectangle
   % BE_8.8.2018: mathforum.org seemed to become https://www.nctm.org/mathforum/.
   }

\begin{figure}[!hb]
\begin{center}
\includegraphics[scale=0.41]{figures/Ulam_spiral_via-CT2-within-1-999.png}
\caption[Primes within the first 999 integers -- as Ulam spiral]
        {Primes within the first 999 integers -- as Ulam spiral\protect\footnotemark}
\label{Primes-in-a-999-integer-ulam-spiral-figure}
\end{center}
\end{figure}
\footnotetext{%
   Graphics from CT2, menu Crypto Tutorials, World of Primes,
   Distribution of primes, Ulam's spiral; 32*32 points.\index{CT2}
}
%\newpage
\begin{figure}[!hb]
\begin{center}
\includegraphics{figures/Ulam_spiral_Wikipedia.png}
\caption[Primes within the first 4000 integers -- as Ulam spiral]
        {Primes within the first 4000 integers -- as Ulam spiral\protect\footnotemark}
\label{Primes-in-a-4000-integer-ulam-spiral-figure}
\end{center}
\end{figure}
\footnotetext{%
   Graphics from 
   \url{http://mathforum.org/mathimages/index.php/Image:Ulam_spiral.png}, 200*200 Ulam spiral
   % BE_8.8.2018: mathforum.org seemed to become https://www.nctm.org/mathforum/.
   }

@@

\hline
\end{tabular}
\caption{The 30+ largest known primes and its particular number types
         (as of Apr 2018)}    % Eyecatcher_neue-Mersenne
\label{L_n_Largest_Known-Primes}
\end{center}
\end{table} 
\addtocounter{footnote}{-1}
\footnotetext{\index{number!Sierpinski}\index{Seventeen or Bust SoB}%
This number was found within the distributed computing project
``Seventeen or Bust'' (SoB)  (\url{https://en.wikipedia.org/wiki/Seventeen_or_Bust})
at March 26, 2007. While the well known \hyperlink{GIMPS-project}{GIMPS project} (chapter~\ref{zahlentyp_mersenne}) searches for bigger and bigger of the infinitely many primes, there is a chance, that the SoB project could have been completed its task sometime.

@@

% --------------------------------------------------------------------------
\vskip +25 pt
\subsection{Challenge of the Electronic Frontier Foundation (EFF)}\index{EFF}
This search is also spurred on by a competition started by the non-profit
organization EFF (Electronic Frontier Foundation) using the means of an 
unknown donor. The participants are rewarded with a total of 500,000 USD if
they find the longest prime number. In promoting this project, the unknown
donor is not looking for the quickest computer, but rather wants to draw 
people's attention to the opportunities offered by {\em cooperative networking} \\
{\url{https://www.eff.org/awards/coop}}

@@

Considerable work has been done on Goldbach's weak conjecture, culminating in a 2013 claim by Harald Helfgott to fully prove the conjecture for all odd integers greater than 7.\footnote{%
Helfgott, H.A. (2013): ``Major arcs for Goldbach's theorem''.
\url{http://arxiv.org/abs/1305.2897}.\\
v4 of his 79 pages paper (14 Apr 2014) states, that the proof works starting at $10^{29}$ (instead of $10^{30})$. And for all numbers up to $10^{30}$ he claims, that he and David Platt proved at the computer that the weak Goldbach is valid.
\\
% \url{http://truthiscool.com/prime-numbers-the-271-year-old-puzzle-resolved}.\\
\url{https://www.newscientist.com/article/dn23535-proof-that-an-infinite-number-of-primes-are-paired.html}.\\
\url{www.spiegel.de/wissenschaft/mensch/beweis-fuer-schwache-goldbachsche-vermutung-a-901111.html}.
}

@@

In May 2013 the results of Yitang Zhang\index{Zhang, Yitang} became
known.\footnote{%
    Erica Klarreich (May 19, 2013): ``Unheralded Mathematician Bridges the
    Prime Gap''\\
    \url{https://www.quantamagazine.org/yitang-zhang-proves-landmark-theorem-in-distribution-of-prime-numbers-20130519/}}
Zhang proved, that there are infinitely many ``cousin primes'', or more
concretely that there is some number $H$ smaller than 70 million such that there
are inifinitely many pairs of primes that differ by $H$.\footnote{%
Whereas the gap between the primes of a twin prime is exactly $2$, cousin
primes\index{cousin prime} do denote two primes, which have a gap between them,
which has a value of a bigger, even, but finite number $H$.
      }${}^,$\footnote{%
This is close to the conjecture stated in 1849 by the French mathematician
Alphonse de Polignac that there are infinitely many prime pairs for any
possible even finite gap, not just $2$.
       }${}^,$\footnote{%
In the meantime this minimal gap $H$ of 70 millions was improved in further
work. The according progress is documented in the Polymath8 project (massively
collaborative online mathematical projects): ``Bounded gaps between primes''.
The best known value of $H$ was $4680$ (as of August 2013) and is till now
(as of April 2014) $2460$ -- this is a good progress compared to 70 million,
but far away from $2$.\\
See \url{http://michaelnielsen.org/polymath1/index.php?title=Bounded_gaps_between_primes}
   }

@@

\vskip +25 pt
\hypertarget{HT-GoogleRecruitment2004}{}
\subsubsection{Recruitment at Google in 2004}%
% \paragraph*{Recruitment at Google in 2004}%
\index{Google!recruitment}%
\label{HT-GoogleRecruitment2004}%
\mbox{}
In summer 2004 the company Google used the number $e$ to attract potential
employees.\footnote{The base of the natural logarithm\index{logarithm!natural}
$e$ is approximately 2.718 281 828 459.
This is one of the most important numbers in all of mathematics like complex
analysis, finance, physics and geometry. Now it was used the first time -- as
far as I know -- for marketing or recruitment.}${}^,$\footnote{Most of this
information is taken from the article ``e-number crunching'' by John Allen
Paulos in TheGuardian, Sept. 30, 2004, and from the web:\\
- \url{https://mkaz.blog/math/google-billboard-problems/}\\
- \url{http://epramono.blogspot.com/2004/10/7427466391.html}\\
- \url{http://mathworld.wolfram.com/news/2004-10-13/google/}.
}
